# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

We load the data into our environment:

```{r}
data<-read.csv("activity.csv")
```
 
Let's take a look at the data:

```{r}
str(data)
```
 
We transform the date column into date-types:

```{r}
data$date<-as.Date(as.character(data$date))
```
 
And check the data again, to verify that our transformation worked:

```{r}
str(data)
```


## What is mean total number of steps taken per day?

We sum the number of steps per day and plot the histogram:

```{r, chunk="histogram"}
library(plyr)
spd<-ddply(data, .(date), summarize, steps_per_day = sum(steps))
spd_cc<-na.omit(spd)
par(mar=c(4,5,1.5,1.5))
with(spd_cc, plot(date,steps_per_day, type="h", lwd=5, col="gold", main="Total number of steps per day", ylab="Steps per day", las=1,mgp=c(3,0.2,0), tck=0.02))
 
```

Calculation of the mean and median total number of steps per day:

```{r}
df<-ddply(data, .(date), summarize,mean_value = mean(steps, na.rm=TRUE), 
          median_value = median(steps, na.rm=TRUE))
print(df)
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
